Denotational Semantics of General Payment Primitives, and Its Payment System

### 2022-10-03 packages/spec-haskell/yellowpaper/Paper1.tex
final words chapter DONE

In Chapter 5, some possible further investigations are included for future study purpose.

@@

Here is a visualization of ``machinery'' of the $\omega$ function:

@@

To finish, here is the illustration of data structures for the denotational payment primitives
implemented with agreement framework:

\begin{figure}[h]
    \centering
    \includegraphics[width=1\textwidth]{../assets/agreement-data-structures.png}
    \caption{Agreement Data Structures}
    \label{fig:ac-omega}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Future Investigations}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Restatement in Agda, Correctness and Equivalence Proofs}

\section{Combinator Library for Financial Contracts}

\section{Payment in General Accounting Domains}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Acknowledgments}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

In Chapter 5, some possible further investigations are also included as notes.

@@

Figure \ref{fig:ac-omega} is a visualization of ``machinery'' of the $\omega$ function:

@@

To finish, figure \ref{fig:ac-agreement-data-structures} is the illustration of data structures for
the denotational payment primitives implemented with agreement framework:

\begin{figure}[p]
    \centering
    \includegraphics[width=0.9\textwidth]{../assets/agreement-data-structures.png}
    \caption{Agreement Data Structures}
    \label{fig:ac-agreement-data-structures}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Final Words}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Notes on Future Investigations}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Restatement in Agda, Correctness and Equivalence Proofs}

One of the advantages of using Haskell language was chosen to write the reference implementation is
that it is a language with industrial strength, and the implementation could be easily integrated
into a production code base without much performance compromises.

However, a major disadvantage is that it is not equipped with sufficient apparatus for program
correctness proofs. The best it can offer without engaging more experimental Haskell features is to
use \textit{QuickCheck}\footnote{\url{https://wiki.haskell.org/QuickCheck} - Haskell wiki page for
QuickCheck.} to the necessary laws are not violated using randomized test vector.

To amend this deficiency, \textit{Agda programming language} comes as a great candidate for a
different constructive restatement of the formal specification.

It is based on the insight of the deep connection (equivalence/isomorphism) between logic and
dependently typed programming, often called "the Curry–Howard correspondence", as discovered and
developed during the 20th century (explained in \cite by Phillip Wadler as ``Propositions as
Types''). Agda being a dependently typed functional programming language fully embodies this
insight, hence a good candidate for creating provable correct program.

To learn more about Agda, see refer to these footnotes\footnote{\url{https://plfa.github.io/} -
Programming Language Foundations in
Agda.} \footnote{\url{https://wiki.portal.chalmers.se/agda/pmwiki.php} - Agda
Wiki.} \footnote{\url{https://wiki.portal.chalmers.se/agda/Main/Community} - Agda community page.}.

\subsection{Future Papers on Modern Payment System}

The next yellow paper will formalize how the denotational payment primitives can be stitched
together using a powerful combinator library pattern briefly described
in \cite{peyton2000composing}.

The execution environments for these financial contracts are also fertile ground for new ideas, from
deterministic execution to avoid the usage of buffer, to using distributed ledger technology to
deep-embed these financial contracts in the consensus layer.

\subsection{General Accounting Domains \& Real Time Finance}

\begin{displayquote}
Accounting, also known as accountancy, is the measurement, processing, and communication of
financial and non financial information about economic entities such as businesses and
corporations. \footfullcite{needles2013principles}
\end{displayquote}

The \textit{real time balance} introduced in this paper can be seen as a instance in
the \textit{general accounting domain} specifically for financial contracts in dealing with payment
primitives, a real time version of the cash flow accounting. It is a long with other types,
including balance sheet accounting, income statement accounting. The conversion between these
instances of general accounting domains are often called ``reconciliations''.

More work can be done on how to create a simple and elegant automation system for the
reconciliations.

\section{Conclusions}
%%%%%%%%%%%%%%%%%%%%%

\emph{
We define the term \textbf{real time finance} to mean a financial system where its payment system is
modernized to handle money distribution continuously over time, its financial contracts are
compositional, and its general accounting domain are automated and processed in real time.
}

This paper defines the what constitutes a modern payment system formally specified using
denotational semantics and restated using Haskell programming language.

In future work, we will look into deeper other topics of real time finance, and continue to refine
the methodology used to achieve \textit{simple and precise} specifications guiding correct and
efficient engineering.

\section{Acknowledgments}
%%%%%%%%%%%%%%%%%%%%%%%%%

(To be confirmed.)
